First steps towards a formalization of Forcing

### 2018-06-24 LSFA/conclusion.tex
[itemize -> prose] complete

Other milestones in this formalization project involve the definition
of the forcing relation, proving the Fundamental Theorem of forcing
(that relates truth in $M$ to that in $M[G]$), and using it to show
that $M[G]\models \ZFC$. The theory is extremely modular in the sense
that the last goal does not depend on the proof of the Fundamental
Theorem nor on the definition of the forcing relation. In a next work,
we will obtain the last goal.

\begin{itemize}
\item Write Isar versions of most proofs in the development.
\item Develop an interface between Paulson's relativization results
  and countable models $M$ of $\ZFC$. This will show
  that every such $M$ is closed under well-founded recursion and, in
  particular, that contains names for each of its
  elements. Consequently, $M\sbq M[G]$.
\item Prove that the basic axioms of $\ZFC$ hold in the generic
  extension. 
\item After prototyping the forcing relation, prove the Axiom Scheme
  of Separation (the first that needs to use the machinery of forcing
  nontrivially).
\end{itemize}

>>>>

Other milestones in this formalization project
involve 
\begin{enumerate}
\item the definition
  of the forcing relation, 
\item proving the Fundamental Theorem of forcing
  (that relates truth in $M$ to that in $M[G]$), and 
\item using it to show
  that $M[G]\models \ZFC$. 
\end{enumerate}
The theory is very modular and this is
witnessed by the fact 
that the last goal does not depend on the proof of the Fundamental
Theorem nor on the definition of the forcing relation. Our next task
will be to obtain the last goal enumerated.

To this end, we will develop  an interface between Paulson's
relativization results and countable models $M$ of $\ZFC$. This will show
that every such $M$ is closed under well-founded recursion and, in
particular, that contains names for each of its
elements. Consequently, the proof of  $M\sbq M[G]$ will be
complete. A landmark will be to prove the Axiom Scheme
of Separation (the first that needs to use the machinery of forcing
nontrivially). As a part of the new formalization, we will provide
Isar versions of the longer applicative proofs presented in this work
